---
title: "Categorizing weightlifting exercises"
author: "Michel Mariën"
date: "18 september 2018"
output:
  html_document: default
  pdf_document: default
---
#### Introduction

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral 
Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering 
the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond
to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six 
male participants aged between 20-28 years, with little weight lifting experience. 

The goal of your project is to predict the manner in which they did the exercise. This is the 
"classe" variable in the training set. You may use any of the other variables to predict with. You 
should create a report describing how you built your model, how you used cross validation, what you 
think the expected out of sample error is, and why you made the choices you did. You will also use 
your prediction model to predict 20 different test cases.

#### Setup

```{r setup, include=TRUE,echo = TRUE,warning=FALSE,message=FALSE}

library(ggplot2)
library(GGally)
library(reshape2)
library(AppliedPredictiveModeling)
library(dplyr)
library(MASS)
library(caret)
library(Hmisc)
library(ElemStatLearn)
library(pgmm)
library(rpart)
library(rattle)
library(randomForest)
library(gridExtra)
library(devtools)
library(janitor)
library(tidyr)

setwd("C:\\_Data\\Mijn Documenten\\R\\Data Science specialization\\Scripts Courses\\Course 8 - Week 4")

dataset_train<-read.csv("pml-training.csv",dec=".",fill=TRUE,stringsAsFactors = FALSE)
dataset_test<-read.csv("pml-testing.csv",dec=".",fill=TRUE,stringsAsFactors = FALSE)

data_train<-dataset_train
data_test<-dataset_test

```

#### Tidying dataset

Before any attempt is made to create a machine learning-model, the data is first cleaned. To make
sure all data is numeric after import into R, all columns are converted to numeric, the categories 
and names are converted to factor and the cvtd-columns is converted to POSIXct. 

After that, all columns containing "NA" or are empty are removed from the test and training set.

```{r tidying, include=TRUE,echo = TRUE,warning=FALSE,message=FALSE}

  ## Kolommen naar juiste klasses zetten
    data_train[,-c(2:6,160)] <- sapply( data_train[,-c(2:6,160)], as.numeric )
    data_train[,2] <- as.factor(data_train[,2])
    data_train[,6]<- as.factor(data_train[,6])
    data_train[,160]<- as.factor(data_train[,160])
    data_train[,5]<- as.POSIXct(data_train[,5],format="%d/%m/%Y %H:%M",tz="GMT")  
  
    data_test[,-c(2:6,160)] <- sapply( data_test[,-c(2:6,160)], as.numeric )
    data_test[,2] <- as.factor(data_test[,2])
    data_test[,6]<- as.factor(data_test[,6])
    data_test[,160]<- as.factor(data_test[,160])
    data_test[,5]<- as.POSIXct(data_test[,5],format="%d/%m/%Y %H:%M",tz="GMT")    
  
    hh<-lapply(data_train,class)
    table(unlist(hh))
  
  ## Verwijderen kolommen met NA-waardes 
    Na_rows<-as.data.frame(sapply(data_train, function(x) sum(is.na(x))))
    Na_rows$Namen<-rownames(Na_rows)
    Na_rows<-Na_rows[Na_rows[,1]==0,]
       
    data_tr<-subset(data_train,select=Na_rows[,2])
    data_te<-subset(data_test,select=Na_rows[c(1:59),2])
    data_te<-data_te%>%mutate(ToPredict=data_test[,160])
    
    data_tr<-data_tr[,-c(1:5)]
    data_te<-data_te[,-c(1:5)]  
```    

#### Exploratory analysis

The dataset with 54 variables is to big to visualize. To get a sense of the kind of data we are
working with, the describe function is applied to the data. The output is explicitly excluded from
this document since it contains a pretty big table. 

```{r exploratory, include=TRUE,echo = TRUE, results='hide',warning=FALSE,message=FALSE}

   describe(data_tr)

```

#### Training

For selecting the best possible model to classify the excersises, 4 different models will be
trained and, depending on their accuracy and Kappa-values, the best will be chosen.

```{r training, include=TRUE,echo = TRUE,warning=FALSE,cache=TRUE,message=FALSE,fig.width=10}
    
  ## Create training dataset
    set.seed(555)
    inTrain = createDataPartition(data_tr$classe, p = 0.7,list=FALSE)
    training = data_tr[inTrain,]
    testing = data_tr[-inTrain,]

    fitControl <- trainControl(method = "cv", number=3)

  ## Decision tree
    Model_dtr<-train(classe~.,method="rpart",data=training,trControl=fitControl)
    Model_dtr$finalModel
  ## Random forest    
    Model_rfo<-train(classe~.,method="rf",data=training,trControl=fitControl)
    Model_rfo$finalModel
  ## Linear discriminat analysis
    Model_lda<-train(classe~.,method="lda",data=training,trControl=fitControl)
    Model_lda$finalModel
  ## kNN-model   
    Model_knn<-train(classe~.,method="knn",data=training,trControl=fitControl)
    Model_knn$finalModel

  ## Group results 
    Model_results <- resamples(list(DTREE=Model_dtr, LDA=Model_lda, KNN=Model_knn, RFOREST=Model_rfo))   

    summary(Model_results)
    
    scales<-list(x=list(relation="free"),y=list(relation="free"))
    bwplot(Model_results,scales=scales)

  ## Predict on test-data
    Model_dtr_predict<-predict(Model_dtr, newdata=testing[,-60])
    Model_lda_predict<-predict(Model_lda, newdata=testing[,-60])    
    Model_knn_predict<-predict(Model_knn, newdata=testing[,-60])   
    Model_rfo_predict<-predict(Model_rfo, newdata=testing[,-60])    
    
    cfm_dtr<-confusionMatrix(testing$classe, Model_dtr_predict)
    cfm_lda<-confusionMatrix(testing$classe, Model_lda_predict)
    cfm_knn<-confusionMatrix(testing$classe, Model_knn_predict)
    cfm_rfo<-confusionMatrix(testing$classe, Model_rfo_predict)

    matrix_accu<-rbind(cfm_dtr$overall,cfm_lda$overall,cfm_knn$overall,cfm_rfo$overall)
    rownames(matrix_accu)<-c("Tree","LDA","KNN","RFO")
    matrix_accu 
    
    OoS_error<-mean(testing$classe != Model_rfo_predict) 
    OoS_error
```    
 
From the code above, one can see that the random forest-model has the highest accuracy (0.997) and
Kappa-value (0.997). The Out-of-Sample error for the random forest-model (OoS_error) is 0.0027 which
is very low and should no wrong classifications on the validation set. 
This means that for the final prediction, the "Model_rfo_predict" will be used

#### Prediction on validation-data 

Applying our Random Forest-model on the validation-set
   
```{r prediction, include=TRUE,echo = TRUE,warning=FALSE,message=FALSE}    
    Model_rfo_predict_final<-predict(Model_rfo, newdata=data_test)
    Model_rfo_predict_final
 
```


